Asimilación de distintas observaciones

### 2020-11-19 analysis/chapters/experimento_20181122.Rmd
Analisis de humedad en primeros niveles

```{r q-wind, warning=FALSE, fig.align="center", fig.height=5, fig.width=9, message=FALSE,dev="png"}
fcsts <- expand_grid(fcsts = c("20181122000000", "20181122060000"),
                     exp = c("E4", "E5", "E6", "E7"))

@@

q[bottom_top <= 7, .(q = mean(q), u = u[bottom_top == 1], v = v[bottom_top == 1]), 
  by = .(south_north, west_east, date, exp)] %>% 
  .[coord, on = c("south_north", "west_east")] %>% 
  .[, date := factor(date)] %>% 
  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = q*1000), proj = norargentina_lambert,
                    breaks = seq(0, 0.02, 0.002)*1000) +
  scale_fill_distiller(name = NULL,
                       palette = "BrBG", direction = 1,
                       breaks = seq(0, 0.02, 0.002)*1000,
                       guide = guide_colorstrip(inside = FALSE,
                                                barwidth = 25,
                                                barheight = 0.8)) +
  # geom_vector(aes(x = lon, y = lat,
  #                 dx = u, dy = v), data = function(x) x[is.cross(y, x, 8)],
  #             arrow.angle = 12, size = 0.3) +
  # scale_mag() +
  geom_sf(data = map, inherit.aes = FALSE, color = "black", fill = NA, size = 0.2) +
  coord_sf(xlim = range(coord$lon), ylim = range(coord$lat)) +
  facet_grid(date ~ exp, 
             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
                                         E6 = "SATWND", E7 = "RAD"))) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Low level humidity (g/Kg)",
       x = NULL,
       y = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))
```

@@

@@

```{r}
files <- list.files(here("analysis", "data", "derived_data"), 
                    full.names = TRUE,
                    pattern = "fss_3h")

>>>>

#### Low levels humidity

Specific humidity field averaged over the first 7 levels of the model, from surface to 800 hPa for low topography areas (aproximately).

* AUT add humidity to the model (is correcting a dry model bias or modifing the circulation and transporting humidity from the north?)
* SATWND is not that different from AUT at list in low levels. The humidity is distributed a little bit different in the domain during 06Z
* RAD shows less humidity that AUT and SATWND but more than CONV. RAD seems to reverse (in part) the effect of incorporating automatic station observations in low levels. This can also be seeing in the first levels the difference between RAD and SATWND.

```{r q-wind, warning=FALSE, fig.align="center", message=FALSE,dev="png"}
fcsts <- expand_grid(fcsts = c("20181121180000", "20181122000000", "20181122060000"),
                     exp = c("E4", "E5", "E6", "E7"))

@@

arg_topo <- metR::GetTopography(-75+360, -50+360, -20, -60, resolution = 1/10)
arg_topo[, lon := metR::ConvertLongitude(lon)]

cordillera <- list(
  ggnewscale::new_scale_fill(),
  geom_contour_fill(data = arg_topo, aes(x = lon, y = lat, 
                                         z = h),
                    breaks = seq(1500, 8000, by = 700)),
  scale_fill_gradient(low = "#E2E6E6", high = "#7E7E7E", guide = "none",
                      oob = scales::squish))

q[bottom_top %between% c(0, 7), .(q = mean(q), u = u[bottom_top == 1], v = v[bottom_top == 1]), 
  by = .(south_north, west_east, date, exp)] %>% 
  .[coord, on = c("south_north", "west_east")] %>% 
  .[, date := factor(date)] %>% 
  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = q*1000), proj = norargentina_lambert,
                    breaks = seq(0, 0.02, 0.002)*1000) +
  scale_fill_distiller(name = NULL,
                       palette = "BrBG", direction = 1,
                       breaks = seq(0, 0.02, 0.002)*1000,
                       guide = guide_colorstrip(inside = FALSE,
                                                barwidth = 25,
                                                barheight = 0.8)) +
  # geom_vector(aes(x = lon, y = lat,
  #                 dx = u, dy = v), data = function(x) x[is.cross(y, x, 8)],
  #             arrow.angle = 12, size = 0.3) +
  # scale_mag() +
  cordillera +
  geom_sf(data = map, inherit.aes = FALSE, color = "black", fill = NA, size = 0.2) +
  coord_sf(xlim = range(coord$lon), ylim = range(coord$lat)) +
  facet_grid(date ~ exp, 
             labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
                                         E6 = "SATWND", E7 = "RAD"))) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Low level humidity (g/Kg)",
       x = NULL,
       y = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))
```

Is the analysis correcting a dry bias on the model or the humidity is being transported into the domain?

The innovation (OmB) for surface humidity observations is mainly positive indicating that the model has always less humidity than the atmosfere (reallity?) throghout the domain for a specific cycle.

@@

exp <- c("E4", "E5", "E6", "E7")

q_field <- purrr::map(exp, function(f) {  
  f
  read_diag_conv(paste0("/home/paola.corrales/datosmunin/EXP/", f,
                        "/ANA/20181121180000/diagfiles/asim_conv_20181121180000.ensmean"), 
                 exp = f, member = "000", variable = c("q"))
}) %>% rbindlist()

q_field[usage.flag == 1 & type %in% c(181, 187) & obs.guess < 0.02] %>% 
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = obs.guess)) +
  scale_color_divergent() +
  geom_mapa() +
  facet_wrap(~exp)
```

The evolution of the mean of the absolute OmB is always positive for all four experiments, AUT and SATWND showing higher values follow by RAD.

```{r}
files <- list.files("/home/paola.corrales/datosmunin/EXP/", full.names = TRUE, recursive = TRUE,
                    pattern = "asim_conv_201811*")
files <- files[!str_detect(files, pattern = "mem")]
files <- files[!str_detect(files, pattern = "training")]
files <- files[!str_detect(files, pattern = "prueba")]
files <- files[!str_detect(files, pattern = "analisis")]
files <- files[!str_detect(files, pattern = "derived_data")]

conv <- purrr::map(files, function(f) {

  meta <- unglue(f, "/home/paola.corrales/datosmunin/EXP//{exp}/ANA/{date}/diagfiles/asim_conv_{date_ana}.ensmean")
  read_diag_conv(f, exp = meta[[1]][["exp"]], member = "000", variable = c("q"))
}) %>% rbindlist()

conv[usage.flag == 1 & type %in% c(181, 187), ":="(mean_omb = mean(abs(obs.guess), na.rm = TRUE),
                                                   n = .N), 
     by = .(date, exp)] %>% 
  ggplot(aes(date, mean_omb)) +
  geom_line(aes(color = exp)) +
  geom_point(aes(color = exp, size = n)) +
  scale_size_area(max_size = 3) +
  scale_color_brewer(palette = "Dark2") +
  labs(y = "Mean of abs(OmB) (g/g)", x = NULL) +
  theme_minimal()
```

```{r}
a <- ReadNetCDF("/home/paola.corrales/datosmunin/EXP/E7/ANA/20181121180000/analysis.ensmean", 
                vars = c(p = "P", "PB", qv = "QVAPOR", 
                         lon = "XLONG", lat = "XLAT")) %>% 
  .[, p := p + PB] %>% 
  .[, c("x", "y") := wrf_project(lon, lat)]

a[bottom_top == 12] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z =  p/100), proj = norargentina_lambert) +
  scale_fill_viridis_b()

exp <- c("E4", "E5", "E6", "E7")

pblh <- purrr::map(exp, function(f) {

  ReadNetCDF(paste0("/home/paola.corrales/datosmunin/EXP/", f, "/ANA/20181121180000/analysis.ensmean"), 
             vars = c("PBLH", lon = "XLONG", lat = "XLAT")) %>% 
    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
    .[, exp := f]

}) %>% 
  rbindlist()

pblh[lat %between% c(-38.5, -20.8) & lon %between% c(-67, -54)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z =  PBLH), proj = norargentina_lambert,
                    breaks = seq(0, 5000, 500)) +
  scale_fill_viridis_b(breaks = seq(0, 5000, 500)) +
  geom_mapa() +
  facet_wrap(~ exp) +
  theme_minimal()

pblh[lat %between% c(-38.5, -20.8) & lon %between% c(-67, -54)] %>% 
  .[, .(mean = mean(PBLH)), by = .(exp, south_north)] %>% 
  ggplot(aes(south_north, mean)) +
  geom_point(aes(color = exp))
```

@@

```{r}
files <- list.files(here("analysis", "data", "derived_data"), 
                    full.names = TRUE,
                    pattern = "fss_6h")
